---
title: "Data_source"
author: "Haowei Ni"
date: "2018/11/19"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(readr)
library(readxl)
<<<<<<< HEAD
library(plotly)
library(maps)

=======
library(maps)
library(ggmap)
>>>>>>> 23921c47affc8a7747626c34299fa78bcf83b5f4
```

Read original data 

```{r input_heart_disease_data}
heart_disease_stratify =  read_csv("data/Heart_Disease_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  rename(mortality_rate = data_value) %>%

  mutate(state = state.name[match(state, state.abb)]) %>% 
  select(-data_source, -geographic_level, -class, -topic, -data_value_footnote, -data_value_footnote_symbol, -topic_id, -location_id ) 

heart_disease = heart_disease_stratify %>% 
  filter(stratification1 == "Overall", stratification2 == "Overall") %>% 
  select(-stratification1, -stratification2, -stratification_category1, -stratification_category2) 

heart_disease$mortality_rate[is.na(heart_disease$mortality_rate)] = 0

heart_disease = 
heart_disease %>% 
  group_by(state) %>% 
  summarise(mortality_rate = mean(mortality_rate))
  
  
```

Add air quality data 

```{r}
airquality_2015 = read_csv("data/airquality.csv") %>%
  janitor::clean_names() %>%
  select(state, pm2_5) %>% 
  group_by(state) %>% 
  summarize(sum(pm2_5))

```

Add obesity data

```{r add_obesity_data}
obesity_data = read_csv("data/National_Obesity_By_State.csv") %>%
  janitor::clean_names() %>%
  rename(state = name) %>%
  rename(obesity_percentage = obesity) %>%
  select(state, obesity_percentage) 
data_with_obesity = left_join(heart_disease, obesity_data)
```

Add stroke data

```{r add_stroke_data}
stroke_data = read_csv("data/Stroke_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(stroke_value=data_value)%>%
  rename(state = location_abbr) %>%
  mutate(state = state.name[match(state, state.abb)])%>%

  select(state,stroke_value) %>% 
  group_by(state) %>% 
  filter(!is.na(stroke_value)) %>% 
  summarize(stroke_value = sum(stroke_value)) 
```



Add income

```{r}
income_data = read_excel("data/income_2015.xlsx", range = "A4:D55") %>%
  janitor::clean_names() %>%
  rename(state = united_states, median_income = x55117, income_standard_error = x253) 
data_with_income = left_join(heart_disease,income_data, by = "state")
```




```{r}
data_income_obesity = left_join(income_data,data_with_obesity, by = "state")


smoking_data = read_csv("data/smoking.csv") %>% 
  filter(YEAR == "2015-2016") %>% 
  mutate(year = 2015) %>% 
  rename(state = LocationDesc) %>% 
  select(-YEAR) %>% 

  select(year, state, Data_Value) %>% 

  select(year, state, Data_Value) %>% 
  rename(tobacco_comsumption = Data_Value) %>% 
  group_by(state) %>% 
  summarise(tobacco_consumption = sum(tobacco_comsumption))


data_income_obesity_smoking = left_join(smoking_data, data_income_obesity, by = "state")

data_income_obesity_smoking_air = left_join(airquality_2015, data_income_obesity_smoking, by = "state")




data_income_obesity_smoking = left_join(smoking_data, data_income_obesity, by = "state")
data_income_obesity_smoking_air = left_join(airquality_2015, data_income_obesity_smoking, by = "state")


final_data = left_join(stroke_data, data_income_obesity_smoking_air, by = "state") 



  

```



Find the association between smoke and heart disease mortality
```{r}
  final_data %>%
  mutate(state = forcats::fct_reorder(factor(state), tobacco_consumption)) %>%
  ggplot(aes(x = mortality_rate, y = tobacco_consumption)) + 
  geom_point(aes(color = state), alpha = .5) +
  labs(
    title = "Tabacco Consumption Accross states"
  ) +
  theme(text = element_text(size = 8), axis.text = element_text(angle = 60, hjust = 1), legend.position = "bottom")
 
lm(mortality_rate~tobacco_consumption, data = final_data) %>%
summary()
  
```

Find the association between income and heart disease mortality

```{r}

final_data_income = 
  final_data %>% 
  mutate(state = forcats::fct_reorder(factor(state), median_income)) 

final_data_income %>% 
  ggplot(aes(x = mortality_rate, y = median_income, color = state)) +
  geom_point() +
  theme(text = element_text(size = 8), axis.text.x = element_text(angle = 60, hjust = 1), legend.position = "bottom") + 

  #Add the title and the name for x and y axis. 
  labs(
    title = "Association between Income and Heart Disease Mortality Rate",
    x = "Mortality Rate",
    y = "median_income"
  )


lm(mortality_rate ~ median_income, data = final_data_income) %>% 
  summary()
```

From the lm result, we can observe that median_income is a very significant variable with a p value of 1.3e-08. This indicates there is a strong association between income and heart disease mortality rate

Find the association between airquality and heart disease mortality

```{r}
## make scatterplot 
final_data %>% 
  mutate(state = fct_reorder(state, mortality_rate)) %>% 
  ggplot(aes(x = mortality_rate, y = pm2.5, color = state)) + 
  geom_point() +
  ggtitle("Airquality VS Mortality Rate ") +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 90, size = 6),
        legend.key.size = unit(0.05, "cm")) + 
  labs(x = "Mortality Rate",
       y = "PM2.5") 
## fit simple linear regression model 
air_regression<-lm(final_data$mortality_rate~final_data$pm2.5) 
summary(air_regression)

  
```

From the scatterplot, we can see that the points are spread randomlly. However, the relationship between pm2.5 and mortality rate is unclear. For the states, with low pm2.5, some of them have low mortality rate and some of them have high mortality rate. After we fit the simple regression model, the p-value for pm2.5 is 0.836, so it is a non-significant variable.  

<<<<<<< HEAD

###Map

```{r}

map_data = final_data %>% 
    mutate(state = tolower(state)) 


states <- map_data("state") %>% 
  rename(state = region)
  
a = left_join(states, map_data, by = "state") %>% 
  mutate(text_label = str_c("Region: ", state, 'Mortality rate: ', mortality_rate) ) 

map_us = ggplot() + 
  geom_polygon(data = states, aes(x=long, y = lat, group = group, fill=state),
               color = "white") + 
  coord_fixed(1.3) + 
  guides(fill=FALSE) +
  geom_label(data = a, aes(x = long, y = lat, label = mortality_rate), 
             size = 2)

ggplotly(map_us)


 plot_ly(x = ~longitude, y = ~latitude, type = "scatter", mode = "markers",
          alpha = 0.5, 
          color = ~price,
          text = ~text_label)

nyc_airbnb %>%
  mutate(text_label = str_c("Price: $", price, '\nRating: ', rating)) %>% 
  plot_ly(x = ~longitude, y = ~latitude, type = "scatter", mode = "markers",
          alpha = 0.5, 
          color = ~price,
          text = ~text_label)
```



```{r}
library(plotly)

a$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>"))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

p <- plot_geo(df, locationmode = 'USA-states') %>%
  add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  ) %>%
  colorbar(title = "Millions USD") %>%
  layout(
    title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)',
    geo = g
  )
p
=======
```{r}
states<- map_data("states") 
>>>>>>> 23921c47affc8a7747626c34299fa78bcf83b5f4
```

