---
title: "Data_source"
author: "Haowei Ni"
date: "2018/11/19"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(readr)
library(readxl)
```

Read original data 

```{r input_heart_disease_data}
heart_disease =  read_csv("data/Heart_Disease_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  mutate(state = state.name[match(state, state.abb)])
```

Add air quality data 

```{r}
airquality_2015 = read_csv("data/airquality.csv") %>%
  janitor::clean_names() %>%
  select(State, County, PM2.5)
  
heart_data = left_join(airquality, heart_disease) 
inner_join(airquality, heart_disease)  
anti_join(airquality, heart_disease) 
anti_join(airquality, heart_disease)
```

Add obesity data

```{r add_obesity_data}
obesity_data = read_csv("data/National_Obesity_By_State.csv") %>%
  janitor::clean_names() %>%
  rename(state = name) %>%
  rename(obesity_percentage = obesity) %>%
  select(state, obesity_percentage) 
data_with_obesity = left_join(heart_disease, obesity_data)
```

Add stroke data

library(tidyverse)
library(data.table)
stroke_data = read_csv("data/Stroke_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(stroke_value=data_value)%>%
  rename(state = location_abbr) %>%
  mutate(state = state.name[match(state, state.abb)])%>%
  select(state,stroke_value)

left_join(heart_disease,stroke_data)
```

Add income

```{r}
income_data = read_excel("data/income_2015.xlsx", range = "A4:D55") %>%
  janitor::clean_names() %>%
  rename(state = united_states, median_income = x55117, income_standard_error = x253) 
data_with_income = left_join(heart_disease,income_data, by = "state")
```


<<<<<<< HEAD
Add smoking
```{r}
smoking_data = read_csv("data/smoking.csv")
```

=======
>>>>>>> 552da2d70c15a845c5a40236a7b1383c413e4ca2
```{r}
data_income_obesity = left_join(income_data,data_with_obesity, by = "state")
```

(from here)


Find the association between average stroke value and average heart disease mortality, and SLR for all counties using wtfdata
```{r}
mydata1=read_csv("data/Heart_Disease_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  rename(mortality_rate = data_value) %>%
  mutate(state = state.name[match(state, state.abb)])%>%
  filter(stratification1 == "Overall", stratification2 == "Overall")%>%
  select(state, location_desc, mortality_rate)
  

mydata2=read_csv("data/Stroke_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  rename(stroke_value = data_value) %>%
  mutate(state = state.name[match(state, state.abb)])%>%
  filter(stratification1 == "Overall", stratification2 == "Overall")%>%
  select(state, location_desc, stroke_value)
                                                                                                                    
comdata = left_join(mydata1,mydata2)  
final_data %>%
  mutate(state = forcats::fct_reorder(state,stroke_value)) %>%
  ggplot(aes(x = mortality_rate, y = stroke_value)) + 
  geom_point(aes(color = state), alpha = .5)
comdata %>%
  mutate(state = forcats::fct_reorder(state,stroke_value)) %>%
  ggplot(aes(x = mortality_rate, y = stroke_value)) + 
  geom_point(aes(color = state), alpha = .5)
stroke_slm=lm(mortality_rate~stroke_value,data=comdata)
summary(stroke_slm)
  

```

find association between heart disease and stroke based on gender and race
```{r}
new_heart_gender_data=read_csv("data/Heart_Disease_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  rename(mortality_rate = data_value) %>%
  rename(gender=stratification1)%>%
  rename(race=stratification2)%>%
  mutate(state = state.name[match(state, state.abb)])%>%
  filter(gender%in%c("Male","Female"))%>%
  filter(race=="Overall")%>%
  select(state,location_desc, mortality_rate,gender)

new_stroke_gender_data=read_csv("data/Stroke_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  rename(stroke_value = data_value) %>%
  rename(gender=stratification1)%>%
  rename(race=stratification2)%>%
  mutate(state = state.name[match(state, state.abb)])%>%
  filter(gender%in%c("Male","Female"))%>%
  filter(race=="Overall")%>%
  select(state, location_desc, stroke_value,gender)
new_gender_data=left_join(new_heart_gender_data,new_stroke_gender_data)

new_heart_race_data=read_csv("data/Heart_Disease_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  rename(mortality_rate = data_value) %>%
  rename(gender=stratification1)%>%
  rename(race=stratification2)%>%
  mutate(state = state.name[match(state, state.abb)])%>%
  filter(gender=="Overall")%>%
  filter(!(race=="Overall"))%>%
  select(state,location_desc, mortality_rate,race)

new_stroke_race_data=read_csv("data/Stroke_Mortality_Data_Among_US_Adults__35___by_State_Territory_and_County.csv") %>%
  janitor::clean_names() %>%
  rename(state = location_abbr) %>%
  rename(stroke_value = data_value) %>%
  rename(gender=stratification1)%>%
  rename(race=stratification2)%>%
  mutate(state = state.name[match(state, state.abb)])%>%
  filter(gender=="Overall")%>%
  filter(!(race=="Overall"))%>%
  select(state, location_desc, stroke_value,race)
new_race_data=left_join(new_heart_race_data,new_stroke_race_data)
```

